### 2019-10-10 paper/methods.tex

\section{Model description}
\subsection{Geometry}
The \gls{SD-TMSR} design model was introduced by the \gls{CAS} during the strategic project ``Future Advanced Nuclear Energy $-$ Thorium-based Molten Salt Reactor System (TMSR)" in 2011 \cite{li_optimization_2018,jiang2012advanced,li2015analysis,li2017model}. The design of \gls{SD-TMSR} is inspired by \gls{MSBR} \cite{robertson_conceptual_1971} after some modification in the geometry to control the positive temperature coefficient in MSBR. The \gls{SD-TMSR} model is described deeply in \cite{li_optimization_2018}.
Figure~\ref{fig:ff} illustrates the quarter-core model configuration of the \gls{SD-TMSR}. The active zone is a right cylinder with height and diameter equal to 460 cm. Assemblies of graphite\footnote{We choose graphite density of $2.3$ g/cm$^3$, to validate our results against results in the literature \cite{li_optimization_2018,nuttin2005potential}.} hexagonal prisms fill the core. The side length of the graphite hexagonal prism was optimized in \cite{li_optimization_2018} and found to be 7.5 cm. The liquid fuel circulates continuously through the fuel channels that pierces the graphite hexagonal prisms. The active zone is divided into two different zones to enhance Th-U breeding performance. The radius of the fuel tubes in the outer zone is 5 cm and the radius of the fuel tubes in the inner zone is 3.5 cm. Moreover, axial and radial reflectors from graphite surround the active zone to maximize neutron flux. The core is surrounded by B${_4}$C cylinder that acts as a shield against heat and neutrons radiation. 
Another Ni-based (hastelloy N) cylinder surrounds the whole core and provides structure and heat protection. The main characteristics of the \gls{SD-TMSR} are listed in Table~\ref{tab:table1}.

\begin{figure} % replace 't' with 'b' to \centering [hbp!]
	\includegraphics[width=\textwidth]{ff.png}
	\caption{The quarter-core model configuration of the SD-TMSR a) Longitudinal view b) Cross-sectional view at the horizontal midplane.}
	\label{fig:ff}
\end{figure}

@@

\subsection{Fuel composition}
The general composition of the liquid fuel salt in this work is 70LiF - 17.5BeF$_2$ - 12.5(HM)F$_4$ mole\%, where HM is the heavy metal (i.e. thorium + different fissile materials).
As previously mentioned, the aim of this paper is to simulate the operation of \gls{SD-TMSR} for a long period of time (60 years) with different initial fissile materials and without any external feed of $^{233}$U. Therefore, five different types of initial fissile materials based on \gls{LEU}, Pu, and \gls{TRU} from LWR SF are investigated as follows:\\
	(1)low-enriched uranium (LEU) (19.79\%);\\
	(2)Pu mixed with LEU (19.79\%);\\
	(3)Pu reactor-grade \cite{marka1993explosive};\\
	(4)transuranic (TRU) elements from LWR SF \cite{de2000scenarios};\\
	(5)and $^{233}$U for comparison purpose.\\

The reactor-grade plutonium and \gls{TRU} vector (\%) are summarized in Table~\ref{tab:table2} and ~\ref{tab:table3}, respectively.

@@

The isotopic compositions of plutonium recovered from the spent fuel composition of commercial LEU \gls{PWR} that has released $33$ $GWd/t$ fission energy and has been cooled for $10$ $years$ before reprocessing \cite{oecd1989probabilistic,marka1993explosive}. As well, the isotopic compositions of \gls{TRU} have been taken from the SF of UOX \gls{PWR} (after one use and without multi-recycling) with $60$ $GWd/t$ burnup, and after $5$ $years$ cooling \cite{de2000scenarios}. 
The molar composition of startup fuel for all five cases is listed in Table~\ref{tab:table4}. Meanwhile, the corresponding initial nuclei inventories with different types of fuel are tabulated in Table~\ref{tab:table5}.

\begin{table}  %[!ht]
	\caption{Composition of startup fuel mole\%}
	\vspace{0.1in}
	\begin{tabularx}{\textwidth}{s s s s s s}
		\hline
		Molecule& \gls{LEU} (19.79\%) & Pu mixed with enriched U (19.79 wt-\%) & Pu reactor-grade & \gls{TRU}& $^{233}$U \\
		\hline
		LiF&70&70&70&70&70\\
		BeF$_2$&17.5&17.5&17.5&17.5&17.5\\
		ThF$_4$&8.25&7.5&10.75&	8.65&12.3		\\
		UF$_4$&4.25&4.75&&&	0.2		\\
		PuF$_3$&&0.25&1.75&&		\\
		TRUF$_3$&&&	&3.85	&\\
		\hline
	\end{tabularx}
	\label{tab:table4}
\end{table}

@@

Simulation of Liquid-fueled Molten Salt Reactor (MSR) systems requires computational software that must support online fuel salt reprocessing and refueling \cite{serp2014molten}. In this work, SERPENT-2 version 2.1.31 beta\footnote{SERPENT-2 is a \gls{3D} continuous energy Monte Carlo neutron transport and burn-up code.} \cite{leppanen2014serpent} is used to simulate the full-core of the SD-TMSR with different types of initial fuel. The extension of SERPENT code accounts for continuous online reprocessing and refueling \cite{aufiero2013extended}. Meanwhile, ENDF-VII.0 cross-section library is adopted for all calculations. The results demonstrate full-core runs of $12.5E+06$ neutron history per depletion step. The full burnup time of the SD-TMSR was 60 years with statistical error in $k_{eff}$ equal to $\pm$ $36$ $pcm$. The online extraction of \gls{FPs} and noble gases  provides many benefits for MSRs. For example, it would reduce the fissile inventory required to achieve criticality and improve the breeding ratio.
Figure~\ref{fig:flow} shows a flow chart of the calculation steps.

@@

As shown in Figure~\ref{fig:flow}, after launched the input file, an advanced matrix exponential solution based on the Chebyshev Rational Approximation Method (CRAM) \cite{isotalo2016improving} used to solve the Bateman equation. Then, the system extracted gaseous \gls{FPs} and other materials (non-dissolved metals, lanthanides, and soluble metals except Pu) in a proper time. This can be done by set the flow rate of gaseous \gls{FPs} and other materials from the fuel to the \texttt{FP-waste tank}\footnote{An imaginary tank used to store the gaseous \gls{FPs} and the other materials (non-dissolved metals, lanthanides, and soluble metals except protactinium).}. Specifically, protactinium was removed from the fuel with a certain flow rate into the external tank, \texttt{pa-decay tank}\footnote{An imaginary tank used to store protactinium extracted from the core.}, to decay and produce $^{233}$U\footnote{The $^{233}$Pa is removed and left to decay into $^{233}$U with $\tau_{1/2}$ $\approx$ $27$ $d$.}. The produced $^{233}$U is used as a fresh fissile fuel and the residual $^{233}$U is the net production of $^{233}$U. The MSR burnup routine provided by SERPENT-2 allows changes the flow rates ($mflow$) of the isotopes during reactor operation \cite{aufiero2013extended}. The mass of the fissile and fertile materials needed to achieve criticality is calculated at the end of the cycle. Then, this mass is added to the core at the beginning of the cycle by a certain flow rate.

\section{Feed and extraction rates}
In the present work, two different feed mechanisms are used. The first mechanism allows continuous feed flow of thorium and $^{233}$U from \texttt{pa-decay tank}. In contrast, the second mechanism adopts continuous feed flow of (Heavy Metal (HM) except for Th) + all or part of $^{233}$U from \texttt{Pa-decay tank}.\\
The fission products act as poisons in the MSRs; they negatively impacting the reactivity. Therefore, \gls{FPs} must be extracted during reactor operation. Consider $T_{r}$ as the time during which the total fuel salt is reprocessed and $dN_{e}$ as the amount of particular element $e$ with inventory $N_{e}$ that the \gls{MSR} extracts during time $dt$; thus \cite{nuttin2005potential}

@@

where $\varepsilon_{e}$ is the removal efficiency. Equation \ref{Equ:1} gives the removal constant $\lambda_{e}$ $[s^{-1}]$ (the rate at which the material is removed), where $\lambda_{e}=\varepsilon_{e}/T_{r}$. The removal constant $\lambda_{e}$ of gaseous and other fission products is precisely calculated and summarized in Table~\ref{tab:table6}.
The effective reprocessing time for the gaseous \gls{FPs} and non-dissolved metals was set to 30 seconds (removal constant $\lambda_{e}$ = $-0.0333$ $s^{-1}$), because such elements must be extract promptly and continuously via He bubbling system. In contrast, extracting the soluble \gls{FPs}, lanthanides, and protactinium can be done by the chemical reprocessing (i.e. fluorination and reduction reaction). Therefore, the system reprocesses a specific amount of fuel salt daily. In the present work, the effective extraction time for soluble \gls{FPs} is $\approx$10.59 days ($\lambda_{e}$ = $-1.092\times10^{-6}$ $s^{-1}$), which is equivalent to 5 m$^3$/d of chemical reprocessing rate \cite{nuttin2005potential,li_optimization_2018}. The effective feed rates of the heavy metals (HM) are changed during reactor operation to conserve the total fuel mass and criticality.

>>>>

\section{Model description}
\subsection{Geometry}
The \gls{SD-TMSR} design model was introduced by the \gls{CAS} as a part of  
the strategic project ``Future Advanced Nuclear Energy $-$ Thorium-based 
Molten Salt Reactor System (TMSR)" in 2011 
\cite{li_optimization_2018,jiang2012advanced,li2015analysis,li2017model}. The 
design of \gls{SD-TMSR} is inspired by \gls{MSBR} 
\cite{robertson_conceptual_1971} after some modification in the geometry to 
control the positive temperature coefficient in MSBR. The \gls{SD-TMSR} core 
geometry was described in details by Li \emph{et al.} 
\cite{li_optimization_2018}.
Figure~\ref{fig:ff} illustrates the quarter-core model configuration of the 
\gls{SD-TMSR}. The active zone is a right cylinder with height and diameter 
equal to 460 cm. Assemblies of graphite\footnote{We choose graphite density of 
$2.3$ g/cm$^3$, to validate our results against results in the literature 
\cite{li_optimization_2018,nuttin2005potential}.} hexagonal prisms fill the 
core. The side length of the graphite hexagonal prism was optimized in 
\cite{li_optimization_2018} and found to be 7.5 cm. The liquid fuel circulates 
continuously through the fuel channels that pierces the graphite hexagonal 
prisms. The core is divided into two different zones to enhance 
Th/$^{233}$U breeding performance. The radius of the fuel channels in the 
outer and inner zone are 5 and 3.5 cm, respectively. The axial and radial 
graphite reflectors surround the core to minimize neutron leakage and 
maximize flux in the core. The reflectors are surrounded by B${_4}$C cylinder 
that acts as a radiation shielding. The \gls{SD-TMSR} pressure vessel holds 
the fuel salt, graphite elements, reflector, shielding, intermediate heat 
exchanger (IHX) and made of Ni-based (hastelloy N) alloy. The main 
characteristics of the \gls{SD-TMSR} are listed in Table~\ref{tab:table1}.

\begin{figure} % replace 't' with 'b' to \centering [hbp!]
	\includegraphics[width=\textwidth]{ff.png}
	\caption{$XZ$ (a) and $XY$ (b) section of the quarter-core model of the 
	SD-TMSR.}
	\label{fig:ff}
\end{figure}

@@

\subsection{Fuel composition}
The general composition of the liquid fuel salt in this work is 70LiF - 
17.5BeF$_2$ - 12.5(HM)F$_4$ mole\%, where HM is the heavy metal (misture of 
thorium and other actinides). The aim of this paper is to simulate the  
operation of \gls{SD-TMSR} for 60 years with different startup fissile 
compositions and without any external feed of fissile $^{233}$U which we 
assumed is unavailable. For that reason, five different types of initial 
fissile materials based on \gls{LEU}, Pu, and \gls{TRU} from LWR SF were 
considered:
\begin{enumerate}[label=(\alph*)]
	\item low-enriched uranium (LEU) (19.79\%);
	\item Pu mixed with \gls{LEU} (19.79\%);
	\item Pu reactor-grade \cite{marka1993explosive};
	\item transuranic (TRU) elements from LWR SF \cite{de2000scenarios};
	\item $^{233}$U for comparison purpose \cite{ashraf2019whole_core}.
\end{enumerate}

The reactor-grade plutonium and \gls{TRU} composition are summarized in 
Table~\ref{tab:table2} and ~\ref{tab:table3}, respectively.

@@

For reactor-grade plutonium case, the composition was taken for plutonium 
recovered from the spent fuel composition of commercial \gls{PWR} with the 
average discharge burnup $33$ $GWd/t$ and after $10$ years of cooling before 
reprocessing \cite{oecd1989probabilistic,marka1993explosive}. Similarly, the 
isotopic compositions of \gls{TRU} was taken for the SF of UOX \gls{PWR} 
(after one use, no multi-recycling) with the average discharge $60$ 
$GWd/t$ burnup, and after $5$ years of cooling \cite{de2000scenarios}. 
The molar composition of startup fuel for all five cases is listed in 
Table~\ref{tab:table4}. Meanwhile, the corresponding initial nuclei 
inventories with different types of fuel are summarized in 
Table~\ref{tab:table5}.
\begin{table}  %[!ht]
	\caption{Composition of startup fuel (mole\%)}
	\vspace{0.1in}
	\begin{tabularx}{\textwidth}{p{0.16\textwidth} X p{0.17\textwidth} 
	p{0.14\textwidth} X X}
		\hline
		Fuel salt component& \gls{LEU} (19.79\%) & Pu+enriched U (19.79wt\%) 
		& Pu reactor-grade & \gls{TRU}& $^{233}$U \\
		\hline
		LiF&70&70&70&70&70\\
		BeF$_2$&17.5&17.5&17.5&17.5&17.5\\
		ThF$_4$&8.25&7.5&10.75&	8.65&12.3		\\
		UF$_4$&4.25&4.75&&&	0.2		\\
		PuF$_3$&&0.25&1.75&&		\\
		(TRU)F$_3$&&&	&3.85	&\\
		\hline
	\end{tabularx}
	\label{tab:table4}
\end{table}

@@

Simulation of Liquid-fueled Molten Salt Reactor (MSR) systems requires 
computational software that must support online fuel salt reprocessing and 
refueling \cite{serp2014molten}. In this work, SERPENT-2 version 2.1.31 
beta\footnote{SERPENT-2 is a 3D continuous energy Monte Carlo neutron 
transport and burnup code.} \cite{leppanen2014serpent} is used to simulate the 
full-core of the SD-TMSR with different types of initial fuel. The extension 
of SERPENT accounts for continuous online reprocessing and refueling 
\cite{aufiero2013extended}. ENDF-VII.0 cross-section library was used for all 
calculations in this work. The results demonstrate full-core runs of 
$1.25\times 10^7$ neutron history per depletion step. The full burnup time of 
the SD-TMSR was 60 years with statistical error in $k_{eff}$ equal to $\pm$ 
$36$ $pcm$. The online extraction of \gls{FPs} and other neutron absorbers 
provides many benefits for MSRs. For example, it would reduce the initial 
fissile material inventory required to achieve criticality and improve the 
breeding ratio. Figure~\ref{fig:flow} shows a flow chart of the calculation 
steps.

@@

As shown in Figure~\ref{fig:flow}, after launched the input file, an advanced 
matrix exponential solution based on the Chebyshev Rational Approximation 
Method (CRAM) \cite{isotalo2016improving} used to solve the Bateman equation. 
Then, the system extracted gaseous \gls{FPs} and other materials 
(non-dissolved metals, lanthanides, and soluble metals except Pu) with 
suitable removal rate\footnote{The extraction rate depends on the type of 
poison and its impact on the neutron
economy.}. This can be done by set the 
flow rate of gaseous \gls{FPs} and other materials from the fuel to the 
\texttt{FP-waste tank}\footnote{An imaginary tank used to store the gaseous 
\gls{FPs} and the other materials (non-dissolved metals, lanthanides, and 
soluble metals except protactinium).}. Specifically, protactinium was removed 
from the fuel with a certain flow rate into the external tank, 
\texttt{pa-decay tank}\footnote{An imaginary tank used to store protactinium 
extracted from the core.}, to decay and produce $^{233}$U\footnote{The 
$^{233}$Pa is removed and left to decay into $^{233}$U with $\tau_{1/2}$ 
$\approx$ $27$ $d$.}. The produced $^{233}$U is used as a fresh fissile fuel 
and the residual $^{233}$U is the net production of $^{233}$U. The MSR burnup 
routine provided by SERPENT-2 allows changes the flow rates ($mflow$) of the 
isotopes during reactor operation \cite{aufiero2013extended}. The mass of the 
fissile and fertile materials needed to achieve criticality is calculated at 
the end of the cycle. Then, this mass is added to the core at the beginning of 
the cycle by a certain flow rate.

\section{Feed and extraction rates}
In the present work, two different feed mechanisms are used. The first 
mechanism allows continuous feed flow of thorium from Th stockpile, and 
$^{233}$U from \texttt{pa-decay tank}. In contrast, the second mechanism 
continuously injects Heavy Metal (HM) (excluding Th) and simultaneously feeds  
all or part of produced $^{233}$U from \texttt{Pa-decay tank}. The fission 
products act as poisons in the MSRs; they negatively impacting the reactivity. 
Therefore, \gls{FPs} must be extracted during reactor operation. Consider 
$T_{r}$ as the time during which the total fuel salt is reprocessed and 
$dN_{e}$ as the amount of particular element $e$ with inventory $N_{e}$ that 
the \gls{MSR} extracts during time $dt$; thus \cite{nuttin2005potential}

@@

where $\varepsilon_{e}$ is the removal efficiency. Equation \ref{Equ:1} gives 
the removal constant $\lambda_{e}$ $[s^{-1}]$ (the rate at which the material 
is removed), where $\lambda_{e}=\varepsilon_{e}/T_{r}$. The removal constant 
$\lambda_{e}$ of gaseous and other fission products is precisely calculated 
and summarized in Table~\ref{tab:table6}. The effective reprocessing time for 
the gaseous \gls{FPs} and non-dissolved metals was set to 30 seconds (removal 
constant $\lambda_{e}$ = $-0.0333$ $s^{-1}$), because such elements must be 
extracted promptly and continuously via gas removal system. In contrast, 
extracting the soluble \gls{FPs}, lanthanides, and protactinium can be done by 
the chemical reprocessing (i.e. fluorination and reduction reaction). 
Therefore, the system reprocesses a specific amount of fuel salt daily. In the 
present work, the effective extraction time for soluble \gls{FPs} is 
$\approx$10.59 days ($\lambda_{e}$ = $-1.092\times10^{-6}$ $s^{-1}$), which is 
equivalent to 5 m$^3$/d of chemical reprocessing rate 
\cite{nuttin2005potential,li_optimization_2018}. The effective feed rates of 
the heavy metals (HM) are changed during reactor operation to conserve the 
total fuel mass and criticality.
